---
title: "A feast of time&nbsp;series&nbsp;tools"
author: "Rob J Hyndman & Mitchell&nbsp;O'Hara-Wild"
date: "June 2019"
colortheme: monashblue
titlefontsize: 22pt
toc: true
output:
  binb::monash:
    incremental: no
    fig_height: 4.8
    keep_tex: yes
    includes:
      in_header: preamble.tex
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 11)
)
library(tidyverse)
library(lubridate)
library(tsibble)
library(tsibbledata)
library(feasts)
options(digits = 3, width = 60)
global_economy <- global_economy %>%
  select(Year, Country, GDP, Imports, Exports, Population)
tourism <- tourism %>%
  mutate(
    State = recode(State,
                   "Australian Capital Territory" = "ACT",
                   "New South Wales"="NSW",
                   "Northern Territory" = "NT",
                   "Queensland" = "QLD",
                   "South Australia" = "SA",
                   "Tasmania" = "TAS",
                   "Victoria"="VIC",
                   "Western Australia" = "WA"
    )
  )
holidays <- tourism %>%
  filter(Purpose=="Visiting") %>%
  group_by(State) %>%
  summarise(
    Trips = sum(Trips)
  ) %>%
  select(Quarter, State, Trips)
```

# Overview

## Tidyverts packages

\begin{textblock}{3.8}(8,0)\begin{alertblock}{}\Large\textbf{tidyverts.org}\end{alertblock}\end{textblock}

\placefig{1}{1.4}{width=4cm}{tsibble.png}
\placefig{5}{1.4}{width=4cm}{tsibbledata.png}
\placefig{3}{4.85}{width=4cm}{feasts.png}
\placefig{7}{4.85}{width=4cm}{fable.png}

## Overview

\placefig{.2}{1.5}{width=4cm}{feasts.png}

\begin{textblock}{7.7}(4.7,1.5)\large
\begin{block}{\fontsize{20}{20}\sf\bfseries\textcolor{DarkYellow}{F}eature \textcolor{DarkYellow}{E}xtraction \textcolor{DarkYellow}{A}nd \textcolor{DarkYellow}{S}tatistics for \textcolor{DarkYellow}{T}ime \textcolor{DarkYellow}{S}eries}
\begin{itemize}
\item works with tidy temporal data provided by the tsibble package.
\item produces time series features, decompositions, statistical summaries and visualisations.
\end{itemize}
\end{block}
\end{textblock}


# Tsibbles

## `tsibble` objects

\fontsize{10}{11.2}\sf

```{r, echo = TRUE}
global_economy
```

\only<2->{\begin{textblock}{.75}(2.15,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{1.6}(3.28,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Key\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4>{\begin{textblock}{6.7}(5.5,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Measured variables\phantom{dg}\end{alertblock}
\end{textblock}}


## `tsibble` objects

\fontsize{10}{11.3}\sf

```{r, echo = TRUE}
tourism
```

\only<2->{\begin{textblock}{1.1}(2.1,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{3.9}(3.65,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Keys\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4-5>{\begin{textblock}{1.5}(7.95,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Measure\phantom{dg}\end{alertblock}
\end{textblock}}

\only<5>{\begin{textblock}{3}(9,5)
\begin{block}{}\fontsize{10}{10}\sf Domestic visitor nights in thousands by state/region and purpose.\phantom{dg}\end{block}
\end{textblock}}



## Holidays by state

\fontsize{9}{10}\sf

```r
holidays <- tourism %>%
  filter(Purpose=="Holiday") %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
```

```{r, echo=FALSE}
holidays
```


# Graphics

## Time plots

\fontsize{9}{10}\sf

```{r, echo = TRUE}
holidays %>% autoplot(Trips)
```

## Season plots
\fontsize{9}{10}\sf

```{r graphics1, fig.width=4, fig.height=5}
holidays %>% gg_season(Trips)
```

## Graphics
\fontsize{9}{10}\sf

```{r graphics2}
holidays %>% gg_subseries(Trips)
```

## Graphics
\fontsize{9}{10}\sf

```{r graphics4, fig.width=4, fig.height=5}
holidays %>% ACF(difference(Trips, 4)) %>% autoplot()
```

# Decompositions

## Decompositions

The feasts package supports four common time series decomposition methods:

* Classical decomposition
* STL decomposition
* X11 decomposition
* X-13ARIMA-SEATS decomposition

## Decompositions
\fontsize{9}{10}\sf

```{r dcmp-plot}
holidays %>% STL(Trips ~ season(window = "periodic")) %>%
  autoplot()
```

# Features

## Feature extraction and statistics

\fontsize{9}{10}\sf

```{r features}
tourism %>% features(Trips, features_stl)
```

## Feature extraction and statistics
\fontsize{8}{9}\sf

```{r features-plot, fig.height=4.6}
tourism %>% features(Trips, features_stl) %>%
  ggplot(aes(x=trend_strength, y=seasonal_strength_year, col=Purpose)) +
    geom_point() + facet_wrap(vars(State))
```

\only<2->{\begin{textblock}{4.7}(7.8,7.3)
\begin{alertblock}{}\fontsize{10}{10}\sf
\begin{itemize}\tightlist
\item Holidays more seasonal than other travel.
\item WA has strongest trends.
\end{itemize}
\end{alertblock}\end{textblock}}

## Feature extraction and statistics
\fontsize{9}{9}\sf

Find the most seasonal time series:

```{r extreme}
most_seasonal <- tourism %>%
  features(Trips, features_stl) %>%
  filter(seasonal_strength_year == max(seasonal_strength_year))
```

\pause\vspace*{-0.3cm}

```{r extreme2, fig.height=3.}
tourism %>%
  right_join(most_seasonal, by = c("State","Region","Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) + geom_line() +
  facet_grid(vars(State,Region,Purpose))
```

```{r pca, echo=FALSE}
# Save pdf figures
savepdf <- function(file, width=16, height=10)
{
  fname <<- paste("figs/",file,".pdf",sep="")
  pdf(fname, width=width/2.54, height=height/2.54, pointsize=10)
  par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
}
endpdf <- function()
{
  crop::dev.off.crop(fname)
}
# Compute features
tourism_features <- tourism %>%
  features(Trips, list(features_stl,features_acf,features_pacf))
# Compute PCs
pcs <- tourism_features %>% select(-State, -Region, -Purpose) %>%
  prcomp(scale=TRUE) %>% augment(tourism_features)
# Save some PC plots
savepdf("pca1", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point() + theme(aspect.ratio=1)
endpdf()
savepdf("pca2", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=State)) +
  geom_point() + theme(aspect.ratio=1)
endpdf()
savepdf("pca3", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
endpdf()
# Find outliers
outliers <- pcs %>%
  filter(.fittedPC2 == min(.fittedPC2) | .fittedPC1 == max(.fittedPC1))
savepdf("pca4", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1) +
  geom_point(data=outliers, aes(x=.fittedPC1, y=.fittedPC2), col="black", shape=1, size=3)
endpdf()
```

## Feature extraction and statistics
\fontsize{9}{9}\sf

```{r tourismfeatures}
tourism_features <- tourism %>%
  features(Trips, list(features_stl,features_acf,features_pacf))
```

```{r echo=FALSE}
tourism_features
```


## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{2.5}(9.6,2,4)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on STL, ACF and PACF features
\end{alertblock}
\end{textblock}

```{r pcatable}
pcs <- tourism_features %>% select(-State, -Region, -Purpose) %>%
  prcomp(scale=TRUE) %>% augment(tourism_features)
```

```{r echo=FALSE}
pcs
```

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.5}(.2,4)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on STL, ACF and PACF features
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{4}{height=5.4cm, width=10cm}{pca1}
\vspace*{10cm}


## Feature extraction and statistics
\fontsize{9}{9}\sf


\begin{textblock}{3.5}(.2,4)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on STL, ACF and PACF features
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=State)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{4}{height=5.4cm, width=10cm}{pca2}
\vspace*{10cm}


## Feature extraction and statistics
\fontsize{9}{9}\sf


\begin{textblock}{3.5}(.2,4)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on STL, ACF and PACF features
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
```

\only<1>{\placefig{4}{4}{height=5.4cm, width=10cm}{pca3}}
\only<2>{\placefig{4}{4}{height=5.4cm, width=10cm}{pca4}}
\vspace*{10cm}


## Feature extraction and statistics
\fontsize{9}{9}\sf

```{r outliers2, fig.height=3.4}
pcs %>%
  filter(.fittedPC2 == min(.fittedPC2)) %>%
  left_join(tourism, by = c("State", "Region", "Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) +
    geom_line() +
    facet_grid(vars(State,Region,Purpose)) +
    ggtitle("Outlying time series in PC space") +
    theme(legend.position = "none")
```


## Feature extraction and statistics
\fontsize{9}{9}\sf

```{r outliers3, fig.height=3.4}
pcs %>%
  filter(.fittedPC1 == max(.fittedPC1)) %>%
  left_join(tourism, by = c("State", "Region", "Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) +
    geom_line() +
    facet_grid(vars(State,Region,Purpose)) +
    ggtitle("Outlying time series in PC space") +
    theme(legend.position = "none")
```


## Acknowledgements


\begin{block}{}
\fontsize{11}{11}\sf
\centering\begin{tabular}{l@{\hspace*{1cm}}l}
\includegraphics[height=4cm, width=10cm]{mitch2} &
\includegraphics[height=4cm, width=10cm]{earowang} \\
Mitchell O'Hara-Wild &
Earo Wang \\
\end{tabular}
\end{block}

\vspace*{1cm}

\begin{alertblock}{}
\centerline{\textbf{feasts.tidyverts.org}}
\centerline{\textbf{robjhyndman.com}}
\end{alertblock}

\vspace*{10cm}